---
title: "Real Title Goes Here"
author: "Jessica Scarborough"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    number_sections: TRUE
    code_folding: show
---

```{r setup, echo = FALSE, cache=FALSE}
# This is the only chunk where you can use echo = FALSE
library(knitr)
options(max.print="75")
opts_chunk$set(comment=NA) 
opts_knit$set(width=75)
```

# R Packages and Setup {-}

```{r packages, message = FALSE}
library(knitr); library(rmdformats)
library(here); library(janitor); library(magrittr)
library(rms); library(broom)

# other packages as needed can go here

library(tidyverse)
```

# Data Source

This project utilizes data from the PLACES project 2020 release. The data contain county-level estimates of 27 health-related measures (5 on chronic disease-related unhealthy behaviors, 13 on health outcomes, and 9 on preventative service utilization). There are 3,142 counties included in the data set, covering the 50 states and District of Columbia of the United States. These data are provided by the Centers for Disease Control and Prevention (CDC), Division of Population Health, Epidemiology and Surveillance Branch. It is accessible to the general public and can be accessed by clicking [here](https://chronicdata.cdc.gov/500-Cities-Places/PLACES-Local-Data-for-Better-Health-County-Data-20/swc5-untb) [1].  

Using the 2017 and 2018 Behavioral Risk Factor Surveillance System (BRFSS) survey results, researchers utilized multilevel regression and post-stratification (MRP) to produce small area estimates of the 27 measures in each county. This approach is described by Zhang et al. (2014) [2] and Zhang et al. (2015) [3]. BRFSS is an annual telephone survey, which uses a standardized questionnaire to ask participants across the country about their health risk activities. It is sponsored by most divisions of the CDC, along with other federal agencies. These data are collected with the purpose of assisting organizations in the building and targeting of health promotion activities. 

# The Subjects

Each row in this data set represents a county within the United States. 

# Loading and Tidying the Data

## Loading the Raw Data

Ingest your raw data here.

## Cleaning the Data

All cleaning and tidying goes here, leading to a single tibble in R. Be sure to explain what you're doing and why you're doing it. Do not show listings of the data as you clean it. 

## Another Cleaning the Data subheading.

I expect you'll use several subheadings here to help delineate tasks.

# The Tidy Tibble

## Listing the Tibble

Here, you will list the tibble, after all cleaning is complete. Just list the tibble.

## Size and Identifiers

Do what we asked you to do in the instructions.

## Saving the R data set

Now, save your tibble as an R data set (.Rds file, which you'll also provide to us.)

# The Code Book

See the instructions and be sure that your code book includes all necessary elements.

## Defining the Variables

Here, to help you get started, is the example from the instructions.

Variable | Role | Type | Description
-------- | ---- | ---- | -----------------
`subjectID` | identifier | - | character code for subjects
`sysbp` | outcome | quant | Most Recent Systolic Blood Pressure, in mm Hg
`statin` | input | 2-cat | Has a current statin prescription? (Yes or No)


## Numerical Description

Here's where you would run `describe` from `Hmisc`.

# Linear Regression Plans

## My Quantitative Outcome

Follow the instructions.

## My Planned Predictors (Linear Model)

Follow the instructions.

# Logistic Regression Plans

## My Binary Outcome

Follow the instructions.

## My Planned Predictors (Logistic Model)

Follow the instructions.

# Affirmation 

Be sure to include the text provided in the instructions.

Note that for Task 2 (Analyses and Presentation), we will add in sections called:

- Linear Regression Modeling
- Logistic Regression Modeling
- Discussion

before the Affirmation but otherwise leave the template alone.

# References

1. PLACES Project. Centers for Disease Control and Prevention. Accessed [2021/02/16]. https://www.cdc.gov/places
2. Zhang, X., Holt, J. B., Lu, H., Wheaton, A. G., Ford, E. S., Greenlund, K. J., & Croft, J. B. (2014). Multilevel regression and poststratification for small-area estimation of population health outcomes: a case study of chronic obstructive pulmonary disease prevalence using the behavioral risk factor surveillance system. *American journal of epidemiology*, 179(8), 1025-1033.
3. Zhang, X., Holt, J. B., Yun, S., Lu, H., Greenlund, K. J., & Croft, J. B. (2015). Validation of multilevel regression and poststratification methodology for small area estimation of health indicators from the behavioral risk factor surveillance system. *American journal of epidemiology*, 182(2), 127-137.


# Session Information

```{r}
xfun::session_info()
```

